---
title: "Visualising market opportunities with a circular barplot"
author: "Edgar Brea"
date: "15/10/2020"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 85%;
  margin-left: auto;
  margin-right: auto;
}
</style>

The goal with this plot is to show the largest and fastest growing markets for robotics products and services worldwide. The data comes from a leading market research company that focus on emerging technologies. The plot includes market segments and corresponding parent markets on the top. The bars indicate the market size (calculated as the average size in USDBillion for the 2020-22 period), and the circles the market growth (calculated as the average year-over-year growth percentage for the 2020-22 period).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(treemap)
```

```{r import_data, include=FALSE}
data_raw_Cyb <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_Cyb.xlsx")
data_raw_DSS <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_DSS.xlsx")
data_raw_ICV <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_ICV.xlsx")
data_raw_RAS <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_RAS.xlsx")
data_raw <- rbind(data_raw_Cyb, data_raw_DSS, data_raw_ICV, data_raw_RAS)
```

```{r tidy_data, include=FALSE}
#calculate means across 3yr columns
data <- data_raw %>% 
  mutate(mean_size = rowMeans(select(., starts_with("size_million_")), na.rm = T), mean_growth = rowMeans(select(., starts_with("yoy_growth_perc_")), na.rm = T))
data <- data[,c(1,2,4,11,12)]

#compute ranks
data <- data %>%
  group_by(segmentation_type_2) %>%
  mutate(rank_size = dplyr::dense_rank(desc(mean_size)), rank_growth = dplyr::dense_rank(desc(mean_growth))) #dense rank sets same rank to ties, and leaves no gap for the following rank

#filter segments to get top 20 in size and/or growth on each segment type
data_top20 <- data %>%
  group_by(segmentation_type_2) %>%
  filter(rank_size <= 20 | rank_growth <= 20)

#set labels for plot
data$label <- ifelse(data$rank_size <= 1 | data$rank_growth <= 1, paste0(data$market_name, ">", data$market_segment_name), "")
```

A glimpse of the data:

```{r describe_data}
head(data_top20)
```

```{r format_data, include=FALSE}
my_segm_type2 <- "application" #application, user, product

data_to_plot <- data.frame(filter(data_top20, segmentation_type_2 == my_segm_type2))
data_to_plot$market_name <- gsub("Global", "", data_to_plot$market_name) #remove "global" from market names
data_to_plot$market_name <- gsub("Market", "", data_to_plot$market_name) #remove "Market" from market names
data_to_plot$market_name <- gsub("market", "", data_to_plot$market_name) #remove "market" from market names
data_to_plot$market_name <- as.factor(data_to_plot$market_name)
data_to_plot$mean_size <- data_to_plot$mean_size/1000 #make them $Billion rather than $Million

# Set a number of 'empty bar' to add at the end of each group (market), then rbind it to the data to plot
empty_bar <- 1
to_add <- data.frame(matrix(NA, empty_bar*nlevels(data_to_plot$market_name), ncol(data_to_plot)))
colnames(to_add) <- colnames(data_to_plot)
to_add$market_name <- rep(levels(data_to_plot$market_name), each=empty_bar)
data_to_plot <- rbind(data_to_plot, to_add)
data_to_plot <- data_to_plot %>% arrange(market_name)
data_to_plot$id <- seq(1, nrow(data_to_plot)) # add id

# data frame with thw name and the y position of each label
label_data <- data_to_plot
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) / number_of_bar # Substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse(angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# data frame with additional info to plot group (parent markets) labels
base_data <- data_to_plot %>% 
  group_by(market_name) %>% 
  summarize(start = min(id), end = max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title = mean(c(start, end)))

# vectors with 4 breaks of the y axis for size and growth (to plot horizontal/circular lines)
my_y_axis <- c(max(data_to_plot$mean_size, na.rm = T)/4*1, max(data_to_plot$mean_size, na.rm = T)/4*2, max(data_to_plot$mean_size, na.rm = T)/4*3, max(data_to_plot$mean_size, na.rm = T)/4*4)
my_y_axis <- round(my_y_axis)
my_y_axis_sec <- c(max(data_to_plot$mean_growth, na.rm = T)/4*1, max(data_to_plot$mean_growth, na.rm = T)/4*2, max(data_to_plot$mean_growth, na.rm = T)/4*3, max(data_to_plot$mean_growth, na.rm = T)/4*4)
my_y_axis_sec <- round(my_y_axis_sec)

data_to_plot$max_size <- ifelse(!is.na(data_to_plot$market_segment_name), max(data_to_plot$mean_size, na.rm = T), NA)
```

The plot code using ggplot:

```{r plot, message=FALSE, warning=FALSE, fig.width=12, fig.height=12, fig.align="center"}
ggplot(data_to_plot, aes(x = as.factor(id), y = mean_size, fill = market_name)) +
  geom_hline(yintercept = my_y_axis[4], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[3], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[2], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[1], colour = "grey", size = .15) +
  geom_hline(yintercept = 0, colour = "grey", size = .15) +
  geom_bar(aes(x = as.factor(id), y = max_size), fill = "white", colour = "white", stat = "identity") + #dummy bars in white to cut through the axis lines
  geom_bar(aes(x = as.factor(id), y = mean_size, fill = market_name), stat = "identity") + #market size
  geom_point(aes(x = as.factor(id), y = mean_growth/5.5, fill = market_name), colour = "grey30", shape = 21, stroke = 1.25) + #market growth
  scale_y_continuous(sec.axis = sec_axis(trans=~.*5.5)) +
  annotate("text", x = rep(max(data_to_plot$id),4), y = my_y_axis, label = paste0(as.character(my_y_axis), "$B|", as.character(my_y_axis_sec), "%"), color="grey60", size=3 , angle=0, fontface="bold", hjust=0.5) +
  ylim(-20, 35) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    panel.grid = element_blank()) +
  coord_polar() + 
  geom_text(data = label_data, aes(x = id, y = 0.1, label = market_segment_name, hjust = hjust), color = "grey36", fontface = "bold", size = 2.75, angle = label_data$angle, inherit.aes = F) + #market segment labels
  geom_text(data = base_data, aes(x = title, y = 30, label = market_name, colour = market_name), size = 3.25, fontface = "bold", inherit.aes = F) + #market label
  ggtitle(paste0("Top largest and fastest growing market segments and corresponding parent markets (segmented by ", my_segm_type2, ""))
```
