---
title: "Visualising robotics market opportunities with a circular barplot"
author: "Edgar Brea"
date: "15/10/2020"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 85%;
  margin-left: auto;
  margin-right: auto;
}
</style>

This plot compares the market size and growth for top 7 global markets for robotics products and services. The data comes from a leading market research company with a focus on emerging technologies. The plot shows parent markets and corresponding segments.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(treemap)
library(stringr)
```

```{r import_data, include=FALSE}
data_raw_Cyb <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_Cyb.xlsx")
data_raw_DSS <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_DSS.xlsx")
data_raw_ICV <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_ICV.xlsx")
data_raw_RAS <- read_excel("~/OneDrive - The University of Queensland/03 CSIRO CPS Strategic Opportunities/05 Data/Phase 2/Technavio Reports/Technavio market data - R ready/market_segments_RAS.xlsx")
data_raw <- rbind(data_raw_Cyb, data_raw_DSS, data_raw_ICV, data_raw_RAS)
```

```{r tidy_data, include=FALSE}
#calculate means across 3yr columns
data <- data_raw %>% 
  mutate(mean_size = rowMeans(select(., starts_with("size_million_")), na.rm = T), mean_growth = rowMeans(select(., starts_with("yoy_growth_perc_")), na.rm = T))
data <- data[,c(1,2,4,11,12)]

#leave only robot markets:
data <- data %>%
  filter(str_detect(market_name, "Robot"))

#leave only segmentation by application:
data <- data %>%
  filter(segmentation_type_2 == "application")

#count markets' segments:
data <- data %>%
  group_by(market_name) %>%
  mutate(no_segments = n())

#filter markets to get those with 4 segments or more:
data <- data %>%
  filter(no_segments >= 5)

#set data to show:
data_vis <- data %>%
  group_by(market_name) %>%
  summarise("number_of_segments" = max(no_segments), "total_size_USDM" = round(sum(mean_size), 2), "avg_growth_rate_YoY_pct" = round(mean(mean_growth), 2))
data_vis <- data.frame(data_vis)

#set labels for plot
data$market_segment_name <- ifelse(nchar(data$market_segment_name) <= 25, data$market_segment_name, paste0(substr(data$market_segment_name,1,25), "..."))
data$market_name <- str_wrap(data$market_name, width = 15)
```

A summary of the market data used for the plot (parent market level):

```{r describe_data}
kable(data_vis)
```

```{r format_data, include=FALSE}
data_to_plot <- data.frame(data)
data_to_plot$market_name <- gsub("Global", "", data_to_plot$market_name) #remove "global" from market names
data_to_plot$market_name <- gsub("Market", "", data_to_plot$market_name) #remove "Market" from market names
data_to_plot$market_name <- gsub("market", "", data_to_plot$market_name) #remove "market" from market names
data_to_plot$market_name <- as.factor(data_to_plot$market_name)
data_to_plot$mean_size <- data_to_plot$mean_size/1000 #make them $Billion rather than $Million

# Set a number of 'empty bar' to add at the end of each group (market), then rbind it to the data to plot
empty_bar <- 1
to_add <- data.frame(matrix(NA, empty_bar*nlevels(data_to_plot$market_name), ncol(data_to_plot)))
colnames(to_add) <- colnames(data_to_plot)
to_add$market_name <- rep(levels(data_to_plot$market_name), each=empty_bar)
data_to_plot <- rbind(data_to_plot, to_add)
data_to_plot <- data_to_plot %>% arrange(market_name)
data_to_plot$id <- seq(1, nrow(data_to_plot)) # add id

# data frame with thw name and the y position of each label
label_data <- data_to_plot
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) / number_of_bar # Substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse(angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# data frame with additional info to plot group (parent markets) labels
base_data <- data_to_plot %>% 
  group_by(market_name) %>% 
  summarize(start = min(id), end = max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title = mean(c(start, end)))

# vectors with 4 breaks of the y axis for size and growth (to plot horizontal/circular lines)
my_y_axis <- c(max(data_to_plot$mean_size, na.rm = T)/4*1, max(data_to_plot$mean_size, na.rm = T)/4*2, max(data_to_plot$mean_size, na.rm = T)/4*3, max(data_to_plot$mean_size, na.rm = T)/4*4)
my_y_axis <- round(my_y_axis,1)
my_y_axis_sec <- c(max(data_to_plot$mean_growth, na.rm = T)/4*1, max(data_to_plot$mean_growth, na.rm = T)/4*2, max(data_to_plot$mean_growth, na.rm = T)/4*3, max(data_to_plot$mean_growth, na.rm = T)/4*4)
my_y_axis_sec <- round(my_y_axis_sec)

data_to_plot$max_size <- ifelse(!is.na(data_to_plot$market_segment_name), max(data_to_plot$mean_size, na.rm = T), NA)
```

The plot was developed using R's ggplot2 package, and it combines barplots and scatter plots arranged in a polar coordinate system. Below is the R code:

```{r plot, message=FALSE, warning=FALSE, fig.width=12, fig.height=12, fig.align="center"}
ggplot(data_to_plot, aes(x = as.factor(id), y = mean_size, fill = market_name)) +
  geom_hline(yintercept = my_y_axis[4], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[3], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[2], colour = "grey", size = .15) +
  geom_hline(yintercept = my_y_axis[1], colour = "grey", size = .15) +
  geom_hline(yintercept = 0, colour = "grey", size = .15) +
  geom_bar(aes(x = as.factor(id), y = mean_size, fill = market_name), stat = "identity", width = 0.6) + #market size
  geom_point(aes(x = as.factor(id), y = mean_growth/30, fill = market_name), colour = "grey30", shape = 21, size = 2, stroke = .75) + #market growth
  scale_y_continuous(limits = c(0,10), sec.axis = sec_axis(trans = ~.*30)) +
  annotate("text", x = rep(max(data_to_plot$id),4), y = my_y_axis, label = paste0("$", as.character(my_y_axis), "B|", as.character(my_y_axis_sec), "%"), color="grey55", size = 2.5 , angle = 0, fontface="bold", hjust=0.5) +
  ylim(-5, 5) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    panel.grid = element_blank(),
    plot.margin = unit(c(-3,-3,-6,-3), "cm")) +
  coord_polar() + 
  geom_text(data = label_data, aes(x = id, y = 0.1, label = market_segment_name, hjust = hjust), color = "grey25", fontface = "bold", size = 2.75, angle = label_data$angle, inherit.aes = F) + #market segment labels
  geom_segment(data=base_data, aes(x = start, y = -0.25, xend = end, yend = -.25), colour = "black", alpha=0.8, size=0.7 , inherit.aes = FALSE )  +
  geom_text(data = base_data, aes(x = title, y = -1.3, label = market_name, colour = market_name), size = 3.25, fontface = "bold", inherit.aes = F) + #market label
  geom_text(x = 0, y = 5, label = "Top largest and fastest growing market segments for robotics products and services", size = 6) + #title
  geom_text(x = 0, y = 4.25, label = "Average for 2020-22 period. Bars represent market size in US$ Billion. Dots represent year-over-year percentage growth.", size = 3.75) #subtitle
```
Thanks to the plot, we see that, although the **Industrial Robot Sensors** and the **Vision Guided Robotics** markets are the largest markets currently, they are not expected to grow substantially in the coming years. Rather, we expect to see a significant growth in **Collaborative Robots**, **Professional Cleaning Robots** and **Agricultural Robots** markets in the coming years. This suggest a gradual shift from the application of robotics in typical industrial settings characterised by repetitive tasks, to the **application of robotics in collaborative, commercial and outdoor environments characterised by unstructured, more dynamic tasks**, a trend likely to be influenced by **recent developments in machine learning and robot dexterity**.


